---
title: "Business Forecasting for fish production in United States"
author: "Vallabh Sawant, Gaurav Vishwakarma, Rakshit Karkera"
---
```{r}
rm(list = ls())
library(fpp)
library(readxl)
library(TTR)
library(xts)
```
```{r}
FP_US <- read_excel("Fish_Production_UnitedStates.xlsx")
head(FP_US)
summary(FP_US)
```
```{r}
FP_US_TS <- ts(FP_US$`Capture fisheries production (metric tons)`, start = c(1960))
#FP_US_TS
#class(FP_US_TS)
summary(FP_US_TS)
plot(FP_US_TS, xlab = "Years", ylab = "Captured fishes (in tonnes)")
grid(nx=NULL, ny=NULL)
```

#The time series shows trend but no seasionality. Also, in the above plot, we can see a sudden rise and the data then gets stable after 1988. Thus we can ignore the previous years and consider the data form 1988 for our forecast.

```{r}
FP_US_TS_recent = window(FP_US_TS,start = c(1988))
summary(FP_US_TS_recent)
```
#The summary shows that the mean has moved more on the higher side since we chose a different window.

```{r}
plot(FP_US_TS, xlab = "Years", ylab = "Captured fishes (in tonnes)", main = "Complete data Time series")
grid(nx=NULL, ny=NULL)
lines(FP_US_TS_recent, col = "Red")
plot(FP_US_TS_recent, xlab = "Years", ylab = "Captured fishes (in tonnes)", main = "Time series of Resampled data")
grid(nx=NULL, ny=NULL)
```

#Looking at the plot, we can come to a conclusion that the data shows no seasonality. It does show some trend. 

```{r}
boxplot(FP_US_TS_recent)
grid(nx=NA, ny=NULL)
```

#As stated in the summary, the boxplot confirms the mean is around 5000000 with an outlier. The 1st quartile is 4858805 where as 3rd quartile is at 5244569. The median is at 4995418 and the mean is 5040506. 

```{r}
Acf(FP_US_TS_recent)
```

#The Acf shows heavy correlation between the current and the previous point as it is shown in the plot above.

```{r}
pacf(FP_US_TS_recent)
```

#The pacf also shows significant correlation between the current and the previous point. 

```{r}
attributes(FP_US_TS_recent)
```

#Naive start
```{r}
naive_forecast = naive(FP_US_TS_recent,3)
plot(naive_forecast, xlab = "Years", ylab = "Captured fishes (in tonnes)")
grid(nx=NULL, ny=NULL)
```

#The above is the plot of naive model. This model basically considers the previous values and forecasts on the basis of that. Snaive won't be present because the data is not seasonal. So technically, it will work as a naive model.

```{r}
attributes(naive_forecast)
```
```{r}
plot(naive_forecast$residuals, xlab = "Years")
grid(nx=NULL, ny=NULL)
```

#Here we don't see a pattern in the graph. This is a good sign. The residuals are scattered randomly.

```{r}
Acf(naive_forecast$residuals, xlab = "Years")
```

#Acf shows no correlation between the residuals. This is ideal.

```{r}
hist(naive_forecast$residuals)
grid(nx=NA, ny=NULL)
```

#The residuals seem a bit right skewed. Since most of the residuals are around the mean that is 0, it seems good.

```{r}
plot(naive_forecast$fitted,naive_forecast$residuals)
grid(nx=NULL, ny=NULL)
```
```{r}
checkresiduals(naive_forecast)
```
```{r}
plot(naive_forecast$fitted,naive_forecast$residuals)
grid(nx=NULL, ny=NULL)
plot.ts(naive_forecast$fitted,naive_forecast$residuals,xy.labels = FALSE,xy.lines = FALSE)
grid(nx=NULL, ny=NULL)
plot.ts(naive_forecast$x,naive_forecast$residuals,xy.labels = FALSE,xy.lines = FALSE)
grid(nx=NULL, ny=NULL)
```


#Here we don't see a pattern in the graph. This is a good sign. The residuals are scattered randomly.


```{r}
accuracy(naive_forecast)
```
```{r}
forecast(naive_forecast,h=3)
```


#As per Naive, the amount of fish to be captured is every year to suffice the population of US is 493017 metric tons with low of 4483870 metric tons and high of 5378164 metric tons. We can use naive bayes as benchmark for other models.

#Naive end

#SMA start
```{r}
par(mfrow=c(2,2))
SMA3_forecast <- ma(FP_US_TS_recent,order = 3)
plot(SMA3_forecast, col = "Red")
grid(nx=NULL, ny=NULL)
SMA6_forecast <- ma(FP_US_TS_recent,order = 6)
plot(SMA6_forecast, col="Green")
grid(nx=NULL, ny=NULL)
SMA9_forecast <- ma(FP_US_TS_recent,order = 9)
plot(SMA9_forecast, col = "Blue")
grid(nx=NULL, ny=NULL)
SMA12_forecast <- ma(FP_US_TS_recent,order = 12)
plot(SMA12_forecast, col = "Yellow")
grid(nx=NULL, ny=NULL)
par(mfrow=c(1,1))
```


```{r}
plot(FP_US_TS_recent, xlab = "Years", ylab = "Captured fishes (in tonnes)", main="Moving Averages Comparison")
lines(SMA3_forecast,col="Red")
lines(SMA6_forecast,col="Green")
lines(SMA9_forecast,col="Blue")
lines(SMA12_forecast,col="yellow")
legend(x = "bottomleft",legend = c("TS","MA3","MA6","MA9","MA12"), lty = c(1,1,1,1,1), col=c("black","red","green","blue","yellow"))
grid(nx=NULL, ny=NULL)
```

```{r}
SMA_forecast <- forecast(SMA3_forecast,h=6)
SMA_forecast
```

#Our data is dependent on the recent values, thus we have decided to go ahead with SMA with order 3. Since the order is smaller, the weight is larger on the recent period.

```{r}
attributes(SMA_forecast)
```
```{r}
plot(SMA_forecast$residuals, xlab = "Years")
grid(nx=NULL, ny=NULL)
```

#Here we don't see a pattern in the graph. This is a good sign. The residuals are scattered randomly.

```{r}
Acf(SMA_forecast$residuals, xlab = "Years")
```
#Acf shows fairly high correlation between the residuals. This is not ideal.

```{r}
hist(SMA_forecast$residuals)
grid(nx=NA, ny=NULL)
```

#The residuals seem a bit right skewed. The histogram also tells us that most of the residuals are on the left side of the mean.

```{r}
plot(SMA_forecast$fitted,SMA_forecast$residuals)
grid(nx=NULL, ny=NULL)
```
```{r}
checkresiduals(SMA_forecast)
```
```{r}
plot(SMA_forecast$fitted,SMA_forecast$residuals)
grid(nx=NULL, ny=NULL)
plot.ts(SMA_forecast$fitted,SMA_forecast$residuals,xy.labels = FALSE,xy.lines = FALSE)
grid(nx=NULL, ny=NULL)
plot.ts(SMA_forecast$x,SMA_forecast$residuals,xy.labels = FALSE,xy.lines = FALSE)
grid(nx=NULL, ny=NULL)
```


#Here we don't see a pattern in the graph. This is a good sign. The residuals are scattered randomly.


```{r}
accuracy(SMA_forecast)
```
#SMA end

#HW start

```{r}
HoltWinters_TS_recent = HoltWinters(FP_US_TS_recent,beta=FALSE,gamma=FALSE)
plot(HoltWinters_TS_recent, xlab="Years")
grid(nx=NULL, ny=NULL)
```

#We are getting aplha as 0.9999339. This is very close to 1 and states that the current values are heavily depenedent on the most recent values.

```{r}
attributes(HoltWinters_TS_recent)
```
```{r}
HoltWinters_residual <- residuals(HoltWinters(FP_US_TS_recent,beta = FALSE, gamma = FALSE))
plot(HoltWinters_residual, xlab = "Years")
grid(nx=NULL, ny=NULL)
```

#We don't see any pattern between the residuals. 

```{r}
Acf(HoltWinters_residual)
```

#The acf states that there is no significant correlation.

```{r}
hist(HoltWinters_residual)
grid(nx=NULL, ny=NULL)
```

#The errors are also distributed around the mean that is 0. This seems good.

```{r}
plot(HoltWinters_TS_recent$fitted,HoltWinters_residual, xlab = "Years")
grid(nx=NULL, ny=NULL)
```

#The above graph is a comparision of the observed and the factors that affect the graph. Here we see the observed and level the same as there is no seasonality to change it.

```{r}
checkresiduals(HoltWinters_TS_recent)
```
```{r}
plot(HoltWinters_TS_recent$fitted,HoltWinters_TS_recent$residuals, xlab = "Years")
grid(nx=NULL, ny=NULL)
plot.ts(HoltWinters_TS_recent$fitted,HoltWinters_TS_recent$residuals,xy.labels = FALSE,xy.lines = FALSE, xlab = "Years")
grid(nx=NULL, ny=NULL)
plot.ts(HoltWinters_TS_recent$x,HoltWinters_TS_recent$residuals,xy.labels = FALSE,xy.lines = FALSE, xlab = "Years")
grid(nx=NULL, ny=NULL)
```

```{r}
HoltWinters_forecast = forecast(HoltWinters_TS_recent,h=3)
HoltWinters_forecast
plot(HoltWinters_forecast)
grid(nx = NULL, ny = NULL)
```

#As per HoltsWinter, the total number of fishses to be captured is 4931025 metric tons.

```{r}
accuracy(HoltWinters_forecast)
```
#HW end

#Decomp start
```{r}
ets_forecast = ets(FP_US_TS_recent)
#plot(ets_forecast, xlab = "Years", yaxt='n')
plot(ets_forecast, xlab = "Years")
```

#Again the observed and level is the same as there is no seasonlaity to influence the data.

```{r}
attributes(ets_forecast)
```

```{r}
plot(ets_forecast$residuals, xlab = "Years")
grid(nx=NULL, ny=NULL)
```

#The plot is equally spread residuals around the horizontal line without a distinct pattern. This is a good indication it is not a non-linear relationship.

```{r}
Acf(ets_forecast$residuals)
```

#Here, in the Acf, we don't see any significant lines which states that there is no correlation between the errors.


```{r}
hist(ets_forecast$residuals)
grid(nx=NULL, ny=NULL)
```

#The histogram is seems normal with an outlier. This seems good for the model.

```{r}
checkresiduals(ets_forecast)
```
```{r}
plot(ets_forecast$fitted,ets_forecast$residuals)
grid(nx=NULL, ny=NULL)
plot.ts(ets_forecast$fitted,ets_forecast$residuals,xy.labels = FALSE,xy.lines = FALSE)
grid(nx=NULL, ny=NULL)
plot.ts(ets_forecast$x,ets_forecast$residuals,xy.labels = FALSE,xy.lines = FALSE)
grid(nx=NULL, ny=NULL)
```

#Here we don't see any specific pattern here for the residuals.

```{r}
accuracy(ets_forecast)
```

```{r}
forecast(ets_forecast,h=3)
```

#The decomposition model forecasts that 4931029 metric tons of fishes need to be captured to suffice the needs of United states.

#Decomp end

#ARIMA start
```{r}
ndiffs(FP_US_TS_recent)
```

#This gives the d value in the Arima function. It tells you the number of differences that you should take to make the time series stationary.
```{r}
Arima_FP_TS = auto.arima(FP_US_TS_recent, trace = TRUE, stepwise = FALSE)
Arima_FP_TS
```

#It shows that the best model using Arima is (0,1,0) since it has the lowest AIC.
```{r}
Arima_FP_TS = auto.arima(FP_US_TS_recent, trace = TRUE, stepwise = FALSE, ic = c("bic"))
Arima_FP_TS
```

#This is the same Arima method, that best model using Arima is (0,1,0) considering BIC values.

```{r}
attributes(Arima_FP_TS)
```
```{r}
plot(Arima_FP_TS$residuals, xlab = "Years")
grid(nx=NULL, ny=NULL)
```

#The plot is equally spread residuals around the horizontal line without a distinct pattern. This is a good indication it is not a non-linear relationship.
```{r}
Acf(Arima_FP_TS$residuals)
```

#Here, in the Acf, we don't see any significant lines which states that there is no correlation between the errors.


```{r}
hist(Arima_FP_TS$residuals)
grid(nx=NULL, ny=NULL)
```

#The histogram is right skewed. 
```{r}
checkresiduals(Arima_FP_TS)
```
```{r}
plot(Arima_FP_TS$fitted,Arima_FP_TS$residuals)
grid(nx=NULL, ny=NULL)
plot.ts(Arima_FP_TS$fitted,Arima_FP_TS$residuals,xy.labels = FALSE,xy.lines = FALSE)
grid(nx=NULL, ny=NULL)
plot.ts(Arima_FP_TS$x,Arima_FP_TS$residuals,xy.labels = FALSE,xy.lines = FALSE)
grid(nx=NULL, ny=NULL)
```

#Here we don't see any specific pattern here for the residuals.

```{r}
Arima_Forecast = forecast(Arima_FP_TS, h = 3)
Arima_Forecast
```

#The arima model forecasts that 4931017 metric tons of fishes need to be captured to suffice the needs of United states. 

```{r}
plot(Arima_Forecast, xlab = "Years", ylab = "Captured fishes (in tonnes)")
grid(nx=NULL, ny=NULL)
```
```{r}
accuracy(Arima_Forecast)
```
#ARIMA end

#Working on the model

#RWF start
```{r}
rwf_forecast = rwf(FP_US_TS_recent,3)
plot (rwf_forecast,  xlab = "Years", ylab = "Captured fishes (in tonnes)")
grid(nx=NULL, ny=NULL)
```
```{r}
attributes(rwf_forecast)
```
```{r}
plot(rwf_forecast$residuals, xlab ="Years")
grid(nx=NULL, ny=NULL)
```

#Residuals don't show any pattern.

```{r}
Acf(rwf_forecast$residuals)
```

#Acf shows no sginficant correlation.

```{r}
hist(rwf_forecast$residuals)
grid(nx=NULL, ny=NULL)
```
```{r}
checkresiduals(rwf_forecast)
```
```{r}
plot(rwf_forecast$fitted,rwf_forecast$residuals)
grid(nx=NULL, ny=NULL)
plot.ts(rwf_forecast$fitted,rwf_forecast$residuals,xy.labels = FALSE,xy.lines = FALSE)
grid(nx=NULL, ny=NULL)
plot.ts(rwf_forecast$x,rwf_forecast$residuals,xy.labels = FALSE,xy.lines = FALSE)
grid(nx=NULL, ny=NULL)
```
```{r}
accuracy(rwf_forecast)
```
```{r}
forecast(rwf_forecast,h=3)
```

#The rwf model forecasts that 4931017 metric tons of fishes need to be captured to suffice the needs of United states
#RWF end

#Regression start
```{r}
Regression_forecasts = lm(FP_US$`Capture fisheries production (metric tons)`~lag(FP_US$`Capture fisheries production (metric tons)`,+1),data = FP_US)
Regression_forecasts
```
```{r}
summary(Regression_forecasts)
```

#Looking at the summary, it can be seen that if the recent value is considered without any change, the current value will differ by -1.974e-09. For every increase of one metric ton in lag1, the current value will increase by 1.000e+00. The adjusted R-squared value is 1. It says that the variation in the dependent variable is completely dependent on the variation in the independent variable.
```{r}
par(mfrow=c(2,2))
plot(Regression_forecasts)
par(mfrow=c(1,1))
```
```{r}
accuracy(Regression_forecasts)
```
#Regression end


#If you notice, all models predict the same value for all the 3 years respectively. The predicted values may differ from model to model but it is the same within the model. This happens because the data is not seasonal or cyclic. The model will just calculate the average and post the values out.

#Comparisons
```{r}
#naive
naive_accuracy = accuracy(naive_forecast)
naive_accuracy
#Moving Average
SMA_accuracy = accuracy(SMA_forecast)
SMA_accuracy
#Holt-Winters
HoltWinters_accuracy = accuracy(HoltWinters_forecast)
HoltWinters_accuracy
#Decomp
ets_accuracy = accuracy(ets_forecast)
ets_accuracy
#Arima
Arima_accuracy = accuracy(Arima_Forecast)
Arima_accuracy
#Random Walk
rwf_accuracy = accuracy(rwf_forecast)
rwf_accuracy
#Regression
#Regression_accuracy = accuracy(Regression_forecasts)
#Regression_accuracy

MAPE <- 5
best_accuracy = c()

best_accuracy[1] <- naive_accuracy[MAPE]
best_accuracy[2] <- SMA_accuracy[MAPE]
best_accuracy[3] <- HoltWinters_accuracy[MAPE]
best_accuracy[4] <- ets_accuracy[MAPE]
best_accuracy[5] <- Arima_accuracy[MAPE]
best_accuracy[6] <- rwf_accuracy[MAPE]
#best_accuracy[7] <- Regression_accuracy[5]
```
The matrix
```{r}
best_accuracy

best_accuracy_MAPE = min(best_accuracy)
best_accuracy_MAPE
```

#As you have seen earlier in Acf and the prediction models that the current value is heavily dependent on the most recent value. Regression gives us the best accuracy because we are using the t+1 (lag1) as a parameter which has high corelation and dependency. But there could be a possible overfit. The second lowest accuracy is SMA but we did see sginificant correlation between the residauls so we will not choose the model. Thus we go ahead with arima model as it has the lowest accuracy compared to others. Thus we go ahead with the arima forecast that 4931029 metric tons of fishes need to be captured to suffice the needs of United states.